---
title: "Data Products - Week 2 Assignment"
author: Prabin Shakya
output: html_document
---

## Here is my Date: 

``` {r echo = TRUE}
myTimestamp = Sys.time()

```

This file was created on `r myTimestamp`.

## Introduction

World Bank collects and makes available the data pertaining to education from all the countries. The data can be easily filtered and downloaded from their easy to use web interface at http://databank.worldbank.org/data/reports.aspx?Code=undefined&id=c755d342&report_name=EdStats_Indicators_Report&populartype=series. 

For the purpose of this assignment, I have downloaded educational statistic for all the countries from year 2014 to look at the Net Enrollment Rate for lower secondary school for the female population. This data is used to plot a heat map showing the status for each country where data is available. It was found that the data is available for only 88 countries.


## Initialize

``` {r echo = TRUE, message=F, warning=F}

setwd("H:\\_R\\Course_9\\Week2\\Proj\\")

library(maps)
library(leaflet)
library(rworldmap)
library(rgeos)
library(rgdal)
library(countrycode)


```

## Data Loading and Scrubbing

``` {r echo = TRUE, message=F, warning=F}


## Read file downloaded from WorldBank.
eduData <- read.csv("Data\\Education_Data_WorldBank.csv",na.strings = c("..",""))
names(eduData) <- c("Country","Country_Code","Series","Series_Code","Value")


## Data Scrubbing - Filter NAs
eduData <- eduData[!is.na(eduData$Value),]
eduData$Value <- round(eduData$Value)

head(eduData)

```


## Transform Data for Map Viewing

``` {r echo = TRUE, message=F, warning=F}

## Get Latitude and Longitude for Countries
wmap<- getMap(resolution = "high")
centroids <- gCentroid(wmap,byid=TRUE)
mapData <- as.data.frame(centroids)
mapData <- cbind(rownames(mapData),mapData)
rownames(mapData) <- NULL
names(mapData) <- c("Country", "long","lat")
mapData$Country_Code <- countrycode(sourcevar = mapData$Country,origin="country.name", destination="iso3c")

# Combine the two data sets to fill in centroids in education data
eduDataMerged <- merge(x=eduData, y= mapData, by="Country_Code", all.x=TRUE)

# Lets just look at educational statistics for female population
female_Edu_Data <- eduDataMerged[grep(" female",eduDataMerged$Series),]


# Load the contries polygon data
countries <- readOGR("countriesMap/data/countries.geojson")
countryList<-as.data.frame(countries$ISO_A3)
names(countryList)<-c("Country_Code")
countryList$sn<-1:nrow(countryList)


# Plug in education statistics into the contries polegon data.
countryList <- merge (x=countryList,y=female_Edu_Data,by="Country_Code",all.x = TRUE)
countryListRev <- countryList[,c("sn","Country_Code","Value")]
countryListRev <- countryListRev[!duplicated(countryListRev),]
countryListRev <- countryListRev[order(countryListRev$sn),]
countries$edu_Value<-countryListRev$Value


```

## Plot the Map

``` {r echo = TRUE}

# Color Palette for the map
pal <- colorNumeric(
  palette = "Blues",
  domain = countries$edu_Value)

# Icon 
bookIcon = icons(iconUrl = "http://cdn.mysitemyway.com/etc-mysitemyway/icons/legacy-previews/icons/3d-transparent-glass-icons-culture/022062-3d-transparent-glass-icon-culture-book2.png",iconWidth = 15, iconHeight = 15)


# Make a map
map <- leaflet(countries)
map %>%
   addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~pal(countries$edu_Value)) %>%
   addMarkers(data=female_Edu_Data, icon = bookIcon, label = ~as.character(paste(female_Edu_Data$Country.x," -> ", female_Edu_Data$Value,"%"))) %>%
  addLegend(pal = pal, values = countries$edu_Value, opacity = 1)


```







